### 2004-07-09 Documentation/part1.tex
*** empty log message ***

In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}. Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution. \index{Examples} Since it depends on your installed
configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to complile the example files.

@@

\begin{table}[tbp]
\begin{center}
\begin{tabular}{llp{7cm}c}
\hline
Experiment & Filename & Short description & Refs. \\
\hline 
\multicolumn{3}{l}{\underline{Conventional beams:}} \\
??? & & \\[0.1cm]

\multicolumn{3}{l}{\underline{First-generation superbeams:}} \\
\JHFSK\ ($\nu$) & {\tt JHFSK.exp} & JHF (J-PARC) to Super-Kamiokande, neutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\
\JHFSK\ ($\bar\nu$)& {\tt JHFSKanti.exp} & JHF (J-PARC) to Super-Kamiokande, antineutrino running &  \cite{Huber:2002rs} \\
\NUMI\  ($\nu$), OA $9 \, \mathrm{km}$ & {\tt NUMI9.exp} & NuMI with off-axis angle of $9 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, neutrino running & \cite{Huber:2002rs} \\
\NUMI\  ($\bar{\nu}$), OA $9 \, \mathrm{km}$ & {\tt NUMI9anti.exp} & NuMI with off-axis angle of $9 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, antineutrino running & \cite{Huber:2002rs} \\
\NUMI\  ($\nu$), OA $12 \, \mathrm{km}$ & {\tt NUMI12.exp} & NuMI with off-axis angle of $12 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, neutrino running & \cite{Huber:2002rs} \\
\NUMI\  ($\bar{\nu}$), OA $12 \, \mathrm{km}$ & {\tt NUMI12anti.exp} & NuMI with off-axis angle of $12 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, antineutrino running & \cite{Huber:2002rs} \\
\SPL\  ($\nu$) & {\tt SPL.exp} & SPL (CERN), neutrino running &  ??? \\
\SPL\  ($\bar\nu$) & {\tt SPLanti.exp} & SPL (CERN), antineutrino running & ??? \\[0.1cm]

\multicolumn{3}{l}{\underline{Superbeam upgrades:}} \\
\JHFHK\ ($\nu$) & {\tt JHFHK.exp} & JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade, neutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\
\JHFHK\ ($\bar\nu$)& {\tt JHFHKanti.exp} & JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade, antineutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\[0.1cm]

\multicolumn{3}{l}{\underline{Neutrino factories:}} \\
\NuFactI\ & {\tt NuFact.exp} & Initial stage neutrino factory, symmetric operation in both polarities & \cite{Huber:2002mx} \\
\NuFactII\  & {\tt NuFact2.exp} & Advanced stage neutrino factory, symmetric operation in both polarities & \cite{Huber:2002mx,Huber:2003ak} \\[0.1cm]

\multicolumn{3}{l}{\underline{Reactor experiments:}} \\
\ReactorI\ & {\tt Reactor.exp} & Small reactor experiment with identical near and far detectors & \cite{Huber:2003pm} \\
\ReactorII\ & {\tt Reactor2.exp} & Large reactor experiment with identical near and far detectors & \cite{Huber:2003pm} \\[0.1cm]

\multicolumn{3}{l}{\underline{$\beta$-Beams:}} \\
\Beta\ ($\nu$) & {\tt BETA.exp} & $\beta$-Beam, neutrino running & ??? \\
\Beta\ ($\bar\nu$) & {\tt BETAanti.exp} & $\beta$-Beam, antineutrino running & ??? \\
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{Experiment data files (table)} NEEDS TO BE ADJUSTED WHEN PROTOTYPES ARE AVAILABLE! Different pre-defined experiments, their filenames (to be used in {\tt LoadExperiment}), their short description, and the references in which they are defined. Note that all experiments use in their standard configurations one year of running time. Details about the experiment parameters can be obtained with {\tt InfoExperiment}(Experiment number) after they have been loaded.}
\end{table}

@@

The functions in \GLOBES\ for version control are:
\begin{function}
\GLBNS{glbTestReleaseVersion}
{\tt int glbTestReleaseVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older. 
\end{function}
\begin{function}
\GLBNS{glbTestLibraryVersion}
{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older.
Note that the library and \GLOBES\ versions are not the same. 
\end{function}
\begin{function}
\GLBNS{glbTestLibraryVersion}
{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older.
Note that the library and \GLOBES\ versions are not the same. 
\end{function}
\begin{function}
\GLBNS{glbVersionOfExperiment} 
{\tt const char* glbVersionOfExperiment(int experiment)} returns
the version string of the experiment number {\tt experiment}.
The version string points to the experiment structure, which means
that it cannot be altered and does not have to be destroyed by the
user.
\end{function}

@@

\bi
\item
 Baseline, Target mass, Source power, running time
\item
 Matter density profile
\item
 Systematics
\item
 Threshold function, efficiencies etc.
\ei

@@

Here come the filter functions

>>>>

In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}. Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution. \index{Examples} Since it depends on your installed
configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to complile the example files. Make sure that
the data files (\AEDL\ and supporting files) in the {\tt data} 
subdirectory can be found by the
examples. In case of doubt, you can simply copy the necessary files
into the working directory where the examples are executed.

We will in this part not go into details of the experiment
definition. The pre-defined experiment prototypes in the {\tt data}
subdirectory are summarized in \tabl{experiments}. They correspond
(except from minor modifications) to the experiments in the
respective references in the table.

@@

\begin{table}[tbp]
\begin{center}
\begin{tabular}{llp{7.5cm}c}
\hline
Experiment & Filename & Short description &  Ref. \\
\hline 
\multicolumn{4}{l}{\underline{Conventional beams:}} \\
MINOS & {\tt MINOS.glb} & MINOS exp., 5~yr running time & \cite{Huber:2004ug}\\
OPERA & {\tt OPERA.glb} & OPERA exp., 5~yr running time & \cite{Huber:2004ug}\\
ICARUS & {\tt ICARUS.glb} & ICARUS exp., 5~yr running time & \cite{Huber:2004ug}\\[0.1cm]

\multicolumn{4}{l}{\underline{First-generation superbeams:}} \\
T2K & {\tt JHFSKnew.glb} & J-PARC to Super-K, 5~yr $\nu$-running &  \cite{Huber:2002rs} \\
 & {\tt JHFSKantinew.glb} & J-PARC to Super-K, 5~yr $\bar{\nu}$-running &  \cite{Huber:2002rs} \\
 & {\tt JHFSKcomb.glb} & Same, but 1.25~yr $\nu$- and 3.75~yr $\bar{\nu}$-running &  \cite{Huber:2002rs} \\[0.1cm]
NO$\nu$A & {\tt NUMI9.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\nu$-running & \cite{Huber:2002rs} \\
& {\tt NUMI9anti.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
& {\tt NUMI9comb.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 1.43~yr $\nu$- and 3.57~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
& {\tt NUMI12.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\nu$-running & \cite{Huber:2002rs} \\
& {\tt NUMI12anti.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
& {\tt NUMI12comb.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 1.43~yr $\nu$- and 3.57~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\[0.1cm]

\multicolumn{4}{l}{\underline{Superbeam upgrade:}} \\
J-PARC-HK & {\tt JHFHKAll.glb} & J-PARC to Hyper-K, 2~yr $\nu$- and 6~yr $\bar{\nu}$-running &  \cite{Huber:2002mx} \\[0.1cm]

\multicolumn{4}{l}{\underline{Neutrino factories:}} \\
\NuFactI\ & {\tt NuFact1.glb} & Initial stage NF, 2$\times$2.5~yr running time (each pol.), $m_{\mathrm{Det}} = 10 \, \mathrm{kt}$,
$P_{\mathrm{Tg}} \simeq 0.75 \, \mathrm{MW}$ & \cite{Huber:2002mx} \\
\NuFactII\  & {\tt NuFact2.glb} & Advanced stage NF, 2$\times$4~yr running time (each pol.), $m_{\mathrm{Det}} = 50 \, \mathrm{kt}$,
$P_{\mathrm{Tg}} \simeq  4 \, \mathrm{MW}$ & \cite{Huber:2002mx} \\[0.1cm]

\multicolumn{4}{l}{\underline{Reactor experiments:}} \\
\ReactorI\ & {\tt Reactor1.glb} & Small reactor exp., $\mathcal{L} = 400 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ & \cite{Huber:2003pm} \\
\ReactorII\ & {\tt Reactor2.glb} & Large reactor exp., $\mathcal{L} = 8\, 000 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ & \cite{Huber:2003pm} \\[0.1cm]

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{Experiment data files (table)} Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

@@

The functions in \GLOBES\ for version control are:
\begin{function}
\GLBNS{glbTestReleaseVersion}
{\tt int glbTestReleaseVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer. 
\end{function}
\begin{function}
\GLBNS{glbTestLibraryVersion}
{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer.
Note that the library and \GLOBES\ versions are not the same. 
\end{function}
\begin{function}
\GLBNS{glbTestLibraryVersion}
{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
the {\tt version} string of the format ``X.Y.Z'' is exactly the
used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer.
Note that the library and \GLOBES\ versions are not the same. 
\end{function}
\begin{function}
\GLBNS{glbVersionOfExperiment} 
{\tt const char* glbVersionOfExperiment(int experiment)} returns
the version string of the experiment number {\tt experiment}.
The version string points to the experiment structure, which means
that it cannot be altered and does not have to be destroyed by the
user.
\end{function}

@@

Using external parameters in \AEDL\ files is a very powerful feature
to change experiment parameters at running time which require that
the experiment be re-initialized. For example, one can change the
energy resolution function or the number of energy bins. However,
in some cases there might be complications, such that the number
of pre- or post-smearing efficiencies does not correspond to the number
of energy bins anymore. Therefore, this feature needs to be 
used with care.

In order to use external parameters in \AEDL\ files, one simply
introduces them. For example, an energy resolution function
\begin{quote}
{\tt
energy(\#EnergyResolution1)< \\
\hspace*{1cm} type = 1 \\
\hspace*{1cm} @sigma\_e = \{ myres ,0,0 \} \\
> \\
}
\end{quote}
might be defined in \AEDL , where the energy resolution is proportional
to {\tt myres} $\times$ energy.

In order to use the user-defined variable, one has to assign it 
with {\tt glbDefineAEDLVariable} {\em before} the experiment is initialized with \GLB{glbInitExperiment}:
\begin{function}
\GLBNS{glbDefineAEDLVariable}
{\tt void glbDefineAEDLVariable(const char* name, double value)}
assigns the value {\tt value} to the \AEDL\ variable {\tt name}.
\end{function}
In our energy resolution example, one could now loop over the
energy resolution such as with
\begin{quote}
{\tt
int i; \\
for(i=5;i<20;i++) \\
\{ \\	 
\hspace*{1cm} glbClearExperimentList(); \\
\hspace*{1cm} glbDefineAEDLVariable("myres",0.01*i); \\
\hspace*{1cm} glbInitExperiment(...); \\
\\
\hspace*{1cm} /* do something */ \\
\}
}
\end{quote}
Note that one does not have do re-initialize the oscillation
parameter vectors every time within the loop as long as the
number of experiments does not change.

In order to clear the external variable stack if one is
excessively using it, one can use
\begin{function}
\GLBNS{glbClearAEDLVariables}
{\tt void glbClearAEDLVariables()}
clears the \AEDL\ variable list.
\end{function}
This function is called automatically upon exit of the program.

@@

The oscillation frequency filters to filter fast oscillations
can also be accessed by the user interface. For details of
the filter functions, we refer to \Sec~\ref{sec:energy} of
the \AEDL\ manual.

In particular, there are a number of functions:
\begin{function}
\GLBNS{glbSetFilterState}
{\tt int glbSetFilterState(int on\_off)} sets the filter state to
on (\GLB{GLB\_ON}) or off (\GLB{GLB\_OFF}).
\end{function}
\begin{function}
\GLBNS{glbGetFilterState}
{\tt int glbGetFilterState()} returns the filter state.
\end{function}
\begin{function}
\GLBNS{glbSetFilterStateInExperiment}
{\tt int glbSetFilterStateInExperiment(int exp, int on\_off)} sets the filter state in experiment {\tt exp} to on (\GLB{GLB\_ON}) or off (\GLB{GLB\_OFF}).
\end{function}
\begin{function}
\GLBNS{glbGetFilterStateInExperiment}
{\tt int glbGetFilterStateInExperiment(int exp)} returns the filter state of experiment {\tt exp}.
\end{function}
Analogously, the filter value can be accessed:
\begin{function}
\GLBNS{glbSetFilter}
{\tt int glbSetFilter(double filter)} sets the filter to the
value {\tt filter}.
\end{function}
\begin{function}
\GLBNS{glbGetFilter}
{\tt double glbGetFilter()} returns the filter value.
\end{function}
\begin{function}
\GLBNS{glbSetFilterInExperiment}
{\tt int glbSetFilterInExperiment(int exp, double filter)} sets the filter  in experiment {\tt exp} to the value {\tt value}.
\end{function}
\begin{function}
\GLBNS{glbGetFilterInExperiment}
{\tt double glbGetFilterInExperiment(int exp)} returns the filter value of experiment {\tt exp}.
\end{function}
The return value of all {\tt Set-} functions is $-1$ if they were not successful.
